Using VariantSpark to learn population structure from genomic profiles

### 2015-10-25 doc/publications/Recomb15/recomb15.tex

The mean error for our \kMeans{} clustering models is approximately XXX. This is similar to the deep-learning approach where they achieved an accuracy of 99\% (two misclassified individuals). Our supervised logistic regression approach, however, improves upon this result.
With this approach, we achieve a mean accuracy of 100\% across five folds, with a standard error of 0. %This is on the above three populations, ASW, CHB and GBR, that belong to the super-populations, AFR, EAS and EUR, respectively.
%As we mention in the introduction, we observed a perfect clustering result when we cluster the entirety of these three super-populations, with an ARI of 1.0. However, when we introduce the fourth super-population into the mix, AMR, the ARI drops to 0.84.
Again we observe the complex multivariate analysis to be inferior to the simple logistic regression.

@@

@@

\subsubsection{Logistic regression is more accurate than \kMeans{} in capturing the full population structure}
In this section we train and test our two methods on the full dataset of 14 populations and assess the accuracy with respect to the accurate classification into the 4 super-populations.
Using 5-fold cross-validation to assess the logistic regression classification results, we observe an average ARI of \(0.93 \pm{} XX\). This is an improvement on our previous result from \kMeans{} clustering the same dataset
where we observed an ARI of \(0.83\).
We also record the precision and recall for each of the groups, and observe that for each classification, the precision and recall for EAS were always 1.0. Therefore each logistic regression model we built was able to classify every EAS individual correctly while not including any false positives in this group.
We consistently observed a relatively low precision for the AMR group (\(0.85\pm{} XX \)).

>>>>

The mean error for our \kMeans clustering models is approximately XXX. This is similar to the deep-learning approach where they achieved an accuracy of 99\% (two misclassified individuals). Our supervised logistic regression approach, however, improves upon this result.
With this approach, we achieve an accuracy of 100\%. This is the average across five folds, with a standard error of 0. This is on the above three populations, ASW, CHB and GBR, that belong to the super-populations, AFR, EAS and EUR, respectively.
As we mention in the introduction, we observed a perfect clustering result when we cluster the entirety of these three super-populations, with an ARI of 1.0. However, when we introduce the fourth super-population into the mix, AMR, the ARI drops to 0.84.

@@

Although we see two of our five runs meet the accuracy of the deep-learning approach, where they achieved an accuracy of 99\% (or ARI of 0.975, with two misclassified individuals), our logistic regression approach demonstrates a consistent improvement.
With this approach, we achieve an accuracy of \(99.997 \pm 0.003\) across five folds.

To reiterate, these results are from the three populations, ASW, CHB and GBR, that belong to the super-populations, AFR, EAS and EUR, respectively.
As we mention in the introduction, we observed a perfect clustering result when we cluster the entirety of these three super-populations (rather than one population from each), with an ARI of 1.0. However, when we introduce the fourth super-population into the mix, AMR, the ARI drops to 0.84.
Given the perfect result from the logistic regression models on the three populations, and the improvement over \kMeans{}, we proceed with logistic regression on the four super-populations.

@@

\subsubsection{Logistic regression is more accurate than \kMeans{} in capturing the full population structure}
In this section we train and test our two methods on the full dataset of 14 populations and assess the accuracy with respect to the accurate classification into the 4 super-populations.
Using 10-fold cross-validation to assess the logistic regression classification results, we observe an average ARI of \(0.93 \pm{} XX\). This is an improvement on our previous result from \kMeans{} clustering the same dataset
where we observed an ARI of \(0.83\).
We also record the precision and recall for each of the groups, and observe that for each classification, and both values are always 1.0 for the EAS super-population. This demonstrates that each logistic regression model was able to classify every EAS individual correctly while not including any false positives in this group.
For the AMR group, however, we consistently observed a relatively low precision (\(0.85\pm{} XX \)). This was generally due to individuals from the EUR super-population being misclassified as AMR. As a result, this led to EUR having a relatively low recall (\(0.9 \pm XX \)).
Although not present in the result of every fold, another observation is that AFR individuals would sometimes be misclassified as AMR. In all ten folds, we didn't observe AFR being classified as EUR, and vice-verse, nor AMR individuals being misclassified as another super-population.
These results demonstrate the diversity, or relative difference of some of the groups.
